Algebraic File Synchronization: Adequacy and Completeness

### 2017-10-02 p2/axioms.tex
New "works" symbol, fixes

So that we can describe the effects of commands independently of filesystems,
let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we investigate sequences of commands with a well-defined order.
\begin{mydef}[Sequences of commands: $\cc$ and $\emptyseq$]
We use ``$\cc$'' to combine commands to form a sequence, or combine sequences to form a longer sequence,
with the meaning that the command on the left is applied first:
\[ \alpha\cc\beta\aFS = \beta(\alpha\aFS). \]
% For example, we will write
% $S=\cbdaa{n}\cc\cbfaa{m}$, or $S'=S\cc\cbba{o}$.
%% monoid, but not free monoid
%% monoid: semigroup with identity element
%% Sequences of commands form a free semigroup
Sequences are also partial endofunctions on filesystems, defined only if all commands succeed.
We write $\sets$ for the set of sequences of commands.
They form a monoid, and, as usual,
$\emptyseq$ denotes its unit element, the empty sequence,
which is defined on all filesystems and by definition leaves all filesystems unchanged.
\end{mydef}

\begin{mydef}[$\Dom{S}$ and $\Img{S}$]
For a sequence of commands $S$, $\Dom{S}$ is the domain of $S$,
which is the set of filesystems $S$
(that is, the commands it contains in the given order)
is defined on, that is, it does not break.

We write $\Img{S}$ to denote the image of the domain of the sequence $S$.
\end{mydef}

@@

@@

\begin{myax}\label{ax_separate_nobreaks}
Commands on incomparable nodes do not break every filesystem:
$\cxynv\cc\czwmv \nequiv \cbrk$ where $n\unrel m$.
\end{myax}

@@

>>>>

So that we can describe the effects of commands independently of filesystems,
let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we investigate sequences of commands with a well-defined order.
\begin{mydef}[Sequences of commands: $\cc$ and $\emptyseq$]
We use ``$\cc$'' to combine commands to form a sequence, or combine sequences to form a longer sequence,
with the meaning that the command on the left is applied first:
\[ \alpha\cc\beta\aFS = \beta(\alpha\aFS). \]
% For example, we will write
% $S=\cbdaa{n}\cc\cbfaa{m}$, or $S'=S\cc\cbba{o}$.
%% monoid, but not free monoid
%% monoid: semigroup with identity element
%% Sequences of commands form a free semigroup
Sequences are also partial endofunctions on filesystems, defined only if all commands it contains succeed in the given order.
We write $\sets$ for the set of sequences of commands.
They form a monoid, and, as usual,
$\emptyseq$ denotes its unit element, the empty sequence,
which is defined on all filesystems and by definition leaves all filesystems unchanged.
\end{mydef}

\begin{mydef}[$\Dom{S}$ and $\Img{S}$]
For a sequence of commands $S$, $\Dom{S}$ is the domain of $S$.
We write $\Img{S}$ to denote the image of the domain of the sequence $S$.
\end{mydef}

@@

\begin{mydef}[$\strext$]
We write $A\strext B$ if $A\eqext B$ and $A\nequiv B$.
In particular, we write $\wrks{A}$ if $A$ is defined on some filesystems.
\end{mydef}

@@

\begin{myax}\label{ax_separate_nobreaks}
Commands on incomparable nodes do not break all filesystems:
$\wrksx{\cxynv\cc\czwmv}$ where $n\unrel m$.
\end{myax}

@@

% TODO \nequiv\cbrk -> \wrks

### 2017-10-02 p2/works-inverse.tex
New "works" symbol, fixes

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will use $A\cc\sqs{B}$ to denote the set of sequences $\{A\cc B\whr B\in\sqs{B}\}$,
and $\Dom{\sqs{A}}$ to denote $\bigcap_{A\in\sqs{A}} \Dom{A}$.
\end{mydef}

@@

As a special case, we write $\worksmnil{\sqs{A}}$ if we know that there is at least one filesystem
all sequences in $\sqs{A}$ are defined on.
We also apply these relations to sequences directly;
e.g. we write
$\worksm{A}{B}$ with the meaning $\worksm{\{A\}}{\{B\}}$.
\end{mydef}

In a sense, $A\worksmeqsign B$ is a weaker counterpart of $A\eqext B$, as the former
only requires that $B$ is defined where $A$ is defined, where the latter also requires
that where they are defined, their effect is the same.
It is easy to see that the following corollary is true:

@@

By definiton, for all objects where it is defined, $a\T\typeeq a$.
% $|\setf|$ is also $1$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, all
propositions proven so far apply to them as well.

@@

% TODO \alpha \neq \cbrk
\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are partial bijections over representative filesystems ($\FS\T$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that representative functions are injective except for cases
where they break a filesystem (they are not defined):
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS\T(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore, as there is a one-to-one correspondence between values and types,
$[\GS\T(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

@@

For the left-to-right part of the proposition,
from $\worksm{\ia}{B\cc\ic}$ we know $\Dom{\ia} \subseteq \Dom{B\cc\ic}$.
As the (representative) $B$ is a bijection between its domain and image,
we can use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc\ia} \subseteq \Dom{\ibi\cc B\cc\ic}$,
but the latter is a subset of $\Dom{\ic}$,
and so $\Dom{\ibi\cc\ia} \subseteq \Dom{\ic}$,
which means that $\worksm{\ibi\cc\ia}{\ic}$.
It is easy to see that $\worksm{\ia}{B\cc\ic} \Leftrightarrow \worksm{\ia}{\{B\}}$.

>>>>

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will use $A\cc\sqs{B}$ to denote the set of sequences $\{A\cc B\whr B\in\sqs{B}\}$,
and $\Dom{\sqs{A}}$ to denote $\bigcap_{A\in\sqs{A}} \Dom{A}$.

We also write $\wrks{\sqs{A}}$ if $\varnothing \subset \Dom{\sqs{A}}$, that is,
if we know that there is at least one filesystem
all sequences in $\sqs{A}$ are defined on.
\end{mydef}

@@

We also apply these relations to sequences directly;
e.g. we write
$\worksm{A}{B}$ with the meaning $\Dom{A} \subseteq \Dom{B}$.
\end{mydef}

In a sense, $A\worksmeqsign B$ is a weaker counterpart of $A\eqext B$, as the former
only requires that $B$ is defined where $A$ is defined, 
while the latter also requires
that where they are defined, their effect is the same.
It is easy to see that the following corollary is true:

@@

By definiton, for all objects where it is defined, $a\T\typeeq a$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, 
all propositions proven so far apply to them as well.

@@

% TODO \alpha \neq \cbrk
\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are partial bijections over representative filesystems ($\FS\T$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that representative functions are injective except for cases
where they are not defined:
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS\T(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore, as there is a one-to-one correspondence between values and types,
$[\GS\T(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

@@

For the left-to-right part of the proposition,
from $\worksm{\ia}{B\cc\ic}$ we know $\Dom{\ia} \subseteq \Dom{B\cc\ic}$.
As the (representative) $B$ is a bijection between its domain and image,
we can use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc\ia} \subseteq \Dom{\ibi\cc B\cc\ic}$,
but the latter is a subset of $\Dom{\ic}$,
and so $\Dom{\ibi\cc\ia} \subseteq \Dom{\ic}$,
which means that $\worksm{\ibi\cc\ia}{\ic}$.
It is easy to see that $\worksm{\ia}{B\cc\ic} \Rightarrow \worksm{\ia}{\{B\}}$.
